\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-09 fully-abstract-comp/popl2019/paper.tex
Lars comments.

This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.

@@

% Return pointers/return seals
The return pointer in the StackToken scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
Both of the capabilities in the pair are of course sealed with the same seal.
% One seal per return point
All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack which would break WBCF.
For similar reasons, we cannot allow return seals to be used to seal closures.
% Don't be stupid
Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal (either directly or indirectly) from memory is leaked.

% Note about them vs us, we do nothing that they couldn't do
In the description of the StackToken calling scheme, it has sometimes been in terms of ``them vs us''.
This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
However, \stktokens{} is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
This means that any component can ensure WBCF and LSE by employing \stktokens{}.
% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose \stktokens{}?

@@

\section{Formulating Security with a Fully Abstract Overlay Semantics}
\label{sec:form-secur-with}
A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
Our proof uses a novel approach we call fully abstract overlay semantics.
In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
An overlay semantics is basically a different view on an existing semantics.
For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
When we have defined the overlay semantics, we prove full-abstractness for a compilation from the overlay semantics to the underlying semantics.
Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
In our case, the compilation even turns out to be the identity function!

>>>>

\todo[inline]{In the below, Lars would like to be more precise about the ``in a general way'' below.
  e.g., by contrasting it to our earlier work where the WBCF and LSE properties were only proved via specific examples.}
This approach expresses what it means to enforce our desirable properties in a general way, and makes it clear that we can support a very general class of programs.
Additionally, formulating security of a calling convention in this way makes it potentially reusable in a larger security proof of a full compiler.
The idea is that security of such a compiler could be verified by proving fully abstract compilation with respect to the well-behaved semantics of the target language, so that the proof can rely on native well-bracketedness and local stack frame encapsulation.
This independent result could then be composed with our result to obtain security of the compiler targeting the real target language, by transitivity of full abstraction.

@@

% Return pointers/return seals
The return pointer in the \stktokens{} scheme is a pair of sealed capabilities where the code part of the pair is the old program counter, and the data part is the stack capability for the local stack frame of the caller.
Both of the capabilities in the pair are of course sealed with the same seal.
% One seal per return point
All call points need to be associated with a unique seal (a return seal) that is only used to seal the capabilities in the return pair for that particular call point.
The return seal is what associates the stack frame on the call stack with a specific call point in a program, so if we allowed return seals to be reused, it would be possible to return to a different call point than the one that caused the stack frame to be pushed on the call stack which would break WBCF.
For similar reasons, we cannot allow return seals to be used to seal closures.
% Don't be stupid
Return seals should never be leaked to adversarial code as this would allow them to unseal the local stack frame of a caller breaking LSE.
This goes for direct leaks, such as leaving a seal in a register or writing it to adversarial memory, as well as indirect leaks, where a capability for reading a return seal (either directly or indirectly) from memory is leaked.

% Note about them vs us, we do nothing that they couldn't do
In the description of the \stktokens{} calling scheme, it has sometimes been in terms of ``them vs us''.
This may have created the impression of an asymmetric calling convention that places a special status on trusted components allowing them to protect themselves against adversaries.
However, \stktokens{} is a modular calling scheme no restriction is put on adversarial components that we do not expect trusted components to meet.
This means that any component can ensure WBCF and LSE by employing \stktokens{}.
% KJAA: ^-- Assuming that there is a global stack address capability, no? Can I "suddenly" begin to impose \stktokens{}?

@@

\section{Formulating Security with a Fully Abstract Overlay Semantics}
\label{sec:form-secur-with}
A central contribution of this paper is to prove that \stktokens{} actually guarantees LSE and WBCF.
Our proof uses a novel approach we call fully abstract overlay semantics.
In this approach, we first define an overlay semantics for \trgcm{} which has LSE and WBCF by construction.
An overlay semantics is basically a different view on an existing semantics.
For instance, it may have new syntactic constructs and have special semantics for these things, but they correspond to something on the underlying machine.
In the overlay semantics we are about to define for \trgcm{}, an explicit stack pointer construct is part of the new view.
The stack pointer construct in the overlay corresponds directly to a linear capability in the underlying semantics, and it will have authority over the part of memory that the overlay views as the stack.
When we have defined the overlay semantics, we prove full-abstraction for a compilation from the overlay semantics to the underlying semantics.
Generally speaking, the compilation from an overlay semantics to an underlying semantics should be straightforward as the overlay semantics merely put a new view on the underlying semantics.
In our case, the compilation even turns out to be the identity function!
